# Exploratory Data Analysis

Author: CW  
Reviewer:

## Introduction

### Prerequisites
```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

## Questions

## Variation

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))
```
```{r}
smaller <- diamonds %>% 
  filter(carat < 3)
```

### 7.3.4 Exercises {-}
1. Explore the distribution of each of the `x`, `y`, and `z` variables in diamonds. What do you learn? Think about a diamond and how you might decide which dimension is the length, width, and depth.
```{r}
# remove false data points
diamonds <- diamonds %>% filter(2 < y & y < 20 & 2 < x & 2 < z & z < 20)
ggplot(diamonds) +
  geom_freqpoly(aes(x = x), binwidth = 0.01)
ggplot(diamonds) +
  geom_freqpoly(aes(x = y), binwidth = 0.01)
ggplot(diamonds) +
  geom_freqpoly(aes(x = z), binwidth = 0.01)
# x and y often share value
ggplot(diamonds) +
  geom_point(aes(x = x, y = y)) +
  geom_point(aes(x = x, y = z), color = "blue") +
  coord_fixed()
```

Seems like `x` and `y` should be length and width, and `z` is depth.

2. Explore the distribution of `price`. Do you discover anything unusual or surprising? (Hint: Carefully think about the `binwidth` and make sure you try a wide range of values.)
```{r}
diamonds %>%
  count(cut_width(price,1000))
# stacking by color
diamonds %>% ggplot() + 
  geom_histogram(aes(x = price,color = color),position = "stack",bins = 100)

ggplot(diamonds) + 
  geom_freqpoly(aes(x=price,color = cut),bins = 1000)+
  coord_cartesian(xlim = c(1000,20000))
```

```{r}
ggplot(diamonds) + 
  geom_freqpoly(aes(x=price,color = color),bins = 1000) +
  coord_cartesian(xlim = c(10000,20000))
```


