# Exploratory Data Analysis

Author: CW  
Status: On-going
Reviewer:

## Introduction

### Prerequisites
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(nycflights13)
```

## Questions

## Variation

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))
```
```{r}
smaller <- diamonds %>% 
  filter(carat < 3)
```

### 7.3.4 Exercises {-}
1. Explore the distribution of each of the `x`, `y`, and `z` variables in diamonds. What do you learn? Think about a diamond and how you might decide which dimension is the length, width, and depth.
```{r}
# remove false data points
diamonds <- diamonds %>% filter(2 < y & y < 20 & 2 < x & 2 < z & z < 20)
ggplot(diamonds) +
  geom_freqpoly(aes(x = x), binwidth = 0.01)
ggplot(diamonds) +
  geom_freqpoly(aes(x = y), binwidth = 0.01)
ggplot(diamonds) +
  geom_freqpoly(aes(x = z), binwidth = 0.01)
# x and y often share value
ggplot(diamonds) +
  geom_point(aes(x = x, y = y)) +
  geom_point(aes(x = x, y = z), color = "blue") +
  coord_fixed()
```

Seems like `x` and `y` should be length and width, and `z` is depth.

2. Explore the distribution of `price`. Do you discover anything unusual or surprising? (Hint: Carefully think about the `binwidth` and make sure you try a wide range of values.)

```{r}
# remove false data points
diamonds <- diamonds %>% filter(2 < y & y < 20 & 2 < x & 2 < z & z < 20)
ggplot(diamonds) + 
  geom_freqpoly(aes(x = price), binwidth = 10) +
  xlim(c(1000, 2000))
```

Somehow we don't have diamonds that are priced around $1500.

3. How many diamonds are 0.99 carat? How many are 1 carat? What do you think is the cause of the difference?
```{r}
diamonds %>% filter(carat == 0.99) %>% count()
diamonds %>% filter(carat == 1) %>% count()
ggplot(diamonds) + 
  geom_histogram(aes(x = carat), binwidth = 0.01) +
  xlim(c(0.97, 1.03))
```
There are much more diamonds with 1 carat. I think it is because psychologically, 1 carat represent a whole new level from 0.99 carat, so for makers, it is little more material for much more value.

4. Compare and contrast `coord_cartesian()` vs `xlim()` or `ylim()` when zooming in on a histogram. What happens if you leave `binwidth` unset? What happens if you try and zoom so only half a bar shows?

```{r}
ggplot(diamonds) + 
  geom_histogram(aes(x = carat)) +
  xlim(c(0.97, 1.035))
ggplot(diamonds) + 
  geom_histogram(aes(x = carat)) +
  coord_cartesian(xlim = c(0.97, 1.035))
ggplot(diamonds) + 
  geom_histogram(aes(x = carat), binwidth = 0.01) +
  xlim(c(0.97, 1.035))
ggplot(diamonds) + 
  geom_histogram(aes(x = carat), binwidth = 0.01) +
  coord_cartesian(xlim = c(0.97, 1.035))
```

`coord_cartesian()` plots and cuts, while `xlim()` cuts and plots. So `xlim()` does not show the half bar.

## Missing values
### Exercises
1. What happens to missing values in a histogram? What happens to missing values in a bar chart? Why is there a difference?  
In a bar chart, `NA` is considered as just another category. In a histogram, `NA` is ignored because the x exis has order.

```{r}
set.seed(0)
df <- tibble(norm = rnorm(100)) %>% mutate(inrange = ifelse(norm > 2, NA, norm))
ggplot(df) +
  geom_histogram(aes(x = inrange))
```

`geom_histogram()` removed rows with `NA` values; 

```{r}
df <- diamonds %>% mutate(cut = as.factor(ifelse(y > 7, NA, cut)))
ggplot(df) + geom_bar(aes(x = cut))
```

Apparently `geom_bar()` doesn't remove `NA`, but rather treat it as another factor or category.

2. What does `na.rm = TRUE` do in `mean()` and `sum()`?   
To ignore `NA`s when calculating mean and sum.

## Covariation
### A categorical and continuous variable
#### Exercises
1. Use what you’ve learned to improve the visualisation of the departure times of cancelled vs. non-cancelled flights.
```{r}
flights %>% 
  mutate(cancelled = is.na(dep_time) | is.na(arr_time)) %>% 
  ggplot() +
  geom_boxplot(aes(x = cancelled, y = dep_time))
flights %>% 
  mutate(cancelled = is.na(dep_time) | is.na(arr_time)) %>% 
  filter(cancelled) %>% 
  select(dep_time)
```

Puzzled by this question: how do we have departure times of cancelled flights?

2. What variable in the diamonds dataset is most important for predicting the price of a diamond? How is that variable correlated with cut? Why does the combination of those two relationships lead to lower quality diamonds being more expensive?
