# Tibbles

Author: Chunji Wang   
Reviewer:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

##

###
```{r}
library(tidyverse)
```

## Creating tibbles

Useful functions:
```{r}
as_tibble(iris)
```

##

### Printing
```{r}
nycflights13::flights %>% 
  print(n = 3, width = Inf)
```

You can see a complete list of options by looking at the package help with `package?tibble`.

### Subsetting
`df %>% .$x` is equal to `df$x`

##

## Exercises

1. How can you tell if an object is a tibble? (Hint: try printing `mtcars`, which is a regular data frame).

```{r}
print(mtcars)
print(as_tibble(mtcars))
```

2. Compare and contrast the following operations on a `data.frame` and equivalent `tibble`. What is different? Why might the default `data.frame` behaviours cause you frustration?

```{r}
df <- data.frame(abc = 1, xyz = "a")
tb <- tibble(abc = 1, xyz = "a")
```
Partial matching is annoying; string as factor is annoying.
```{r}
df$x
tb$x
```
Same operiation `[]` to `data.frame`, but results are categorically different.
```{r}
df[, "xyz"]
df[, c("abc", "xyz")]
```
`tibble` always return a tibble as a result of `[]` operation.
```{r}
tb[, "xyz"]
tb[, c("abc", "xyz")]
```

3. If you have the name of a variable stored in an object, e.g. `var <- "mpg"`, how can you extract the reference variable from a tibble?

```{r}
var <- "mpg"
tb <- tibble(mpg = 5, wheel = 4)
tb %>% .[[var]]
```

4. Practice referring to non-syntactic names in the following data frame by:
    1. Extracting the variable called `1`.
    2. Plotting a scatterplot of `1` vs `2`.
    3. Creating a new column called `3` which is `2` divided by `1`.
    4. Renaming the columns to `one`, `two` and `three`.

```{r}
annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)
annoying %>% .$`1`
ggplot2::ggplot(annoying, aes(x = `1`, y = `2`)) + ggplot2::geom_point()
annoying <- annoying %>% mutate(`3` = `2` / `1`)
print(annoying)
annoying <- annoying %>% rename(one = `1`, two = `2`, three = `3`)
print(annoying)
```

5. What does `tibble::enframe()` do? When might you use it?
```{r}
enframe(1:3)
```


6. What option controls how many additional column names are printed at the footer of a tibble?  
`tibble.max_extra_cols`
