# Relational data

Author: Chunji Wang  
Reviewer:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

##
###
```{r}
library(tidyverse)
library(nycflights13)
```

##
### Exercises
1. Imagine you wanted to draw (approximately) the route each plane flies from its origin to its destination. What variables would you need? What tables would you need to combine?  

We need `tailnum` to identify planes, `origin` and `dest` to identify airports; we need to combine `airports` with `flights` to get geolocation of airports.

2. I forgot to draw the relationship between `weather` and `airports`. What is the relationship and how should it appear in the diagram?

They should be connected via `origin`. 

3. `weather` only contains information for the `origin` (NYC) airports. If it contained weather records for all airports in the USA, what additional relation would it define with flights?

It would define another relationship via `dest`.

4. We know that some days of the year are “special”, and fewer people than usual fly on them. How might you represent that data as a data frame? What would be the primary keys of that table? How would it connect to the existing tables?

A table with `dates` as the primary key, and another column `special` indicating whether this date is special or not. It would connect to `flights` via `year` `month` `day` variables.

## Keys
```{r}
planes %>% count(tailnum)
```

### Exercises
1. Add a surrogate key to `flights`.
```{r}
flights %>% mutate(row_number = row_number())
```

2. Identify the keys in the following datasets
    1. `Lahman::Batting`
    2. `babynames::babynames`
    3. `nasaweather::atmos`
    4. `fueleconomy::vehicles`
    5. `ggplot2::diamonds`
```{r}
Lahman::Batting
babynames::babynames
nasaweather::atmos
fueleconomy::vehicles
ggplot2::diamonds
```

3. Draw a diagram illustrating the connections between the `Batting`, `Master`, and `Salaries` tables in the `Lahman` package. Draw another diagram that shows the relationship between `Master`, `Managers`, `AwardsManagers`.  
How would you characterise the relationship between the `Batting`, `Pitching`, and `Fielding` tables?

## Mutating joins


