# Iteration

Author: chunjiw  
Reviewer:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

```{r}
library(tidyverse)
```

##
##
##
###
###
###
###
### Exercises

1. 

2. What happens if you use `for (nm in names(x))` and `x` has no names? What if only some of the elements are named? What if the names are not unique?
```{r}
x <- 1:3
# if there's no name, there's no loop
for (nm in names(x)) {print(is.null(nm))}
# if only some elements are named, there will be a loop, over names and NAs.
names(x) <- c(NULL, 'a', NULL)
for (nm in names(x)) {print(nm)}
# if names are not unique, the loop will be over non-unique names.
names(x) <- c('a', 'a', 'b')
for (nm in names(x)) {print(nm)}
```

3.
4.

## For loops vs. functionals
## The map functions
###
###
### Exercises

1. Write code that uses one of the map functions to:

    1. Compute the mean of every column in `mtcars`.
```{r}
map_dbl(mtcars, mean)
```
    
    2. Determine the type of each column in nycflights13::flights.
    3. Compute the number of unique values in each column of iris.
```{r}
map_int(iris, n_distinct)
```
    
    4. Generate 10 random normals ...

2. How can you create a single vector that for each column in a data frame indicates whether or not it’s a factor?
```{r}
map_lgl(iris, is.factor)
```

3. What happens when you use the map functions on vectors that aren’t lists? `What does map(1:5, runif)` do? Why?

4. 

