# Many models

Author: Ron 
Reviewer:
  
## Introduction

Note:

1. Use many simple models to better understand complex datasets
2. Use list-columns to store arbitrary data stracture in a data frame
3. Use **broom** package to turn models into tidy data

### Prerequisites
```{r}
library(modelr)
library(tidyverse)
library(gapminder)
```

## gapminder
```{r}
gapminder
```

### Nested data

use `nest()` function to create a dataframe of dataframes which has a column called `data`.

```{r}
by_country <- gapminder %>%
  group_by(country, continent) %>%
  nest()
```
Use `mutate` and `purrr::map` to generate new column, the model itself, which is an S3 object.

```{r}
country_model <- function(df) {
  lm(lifeExp ~ year, data = df)
}
by_country <- by_country %>% 
  mutate(model = map(data, country_model))
by_country
```

### Unnesting
`purrr::map2` can map mulitple inputs simultaneously
```{r}
by_country <- by_country %>%
  mutate(resids = map2(data, model, add_residuals)
         # data here is the column name, so is model.
         )

unnest(by_country,resids)
```
Note that each row is one *data point* in the original `gapminder` but the resid calculated is based on the country-wise model.

### Model quality

Use `broom` package to check model quality metrices. `broom::glance` takes a model as input

```{r}
glance <- by_country %>%
  mutate(glance = map(model, broom::glance)) %>%
  unnest(glance, .drop = TRUE)
# drop the list columns, `data`, `model` and `resid`
```
Note that each `model` gives a set of metrics.

```{r}
bad_fit <- filter(glance, r.squared < 0.25)

gapminder %>% 
  semi_join(bad_fit, by = "country") %>% 
  ggplot(aes(year, lifeExp, colour = country)) +
    geom_line()
```

### 25.2.5 Exercises

1. A linear trend seems to be slightly too simple for the overall trend. Can you do better with a quadratic polynomial? How can you interpret the coefficients of the quadratic? (Hint you might want to transform `year` so that it has mean zero.)

```{r}
modquad <- function(df){
  lm(data = df, lifeExp ~ poly(year,2))
}

quad <- gapminder %>%
  mutate(year = year - mean(year)) %>%
  group_by(country)%>%
  nest()%>%
  mutate(model = purrr::map(data,modquad))

quad <- quad %>%
  mutate(resids = map2(data, model, add_residuals)
         # data here is the column name, so is model.
         )
unnest(quad,resids)
```
Check the residuals
```{r}
unnest(quad, resids) %>%
  ggplot(aes(group = country))+
  geom_line(aes(x = year, y = resid))+
  facet_wrap(~continent, nrow = 2)
```
Check the quality
```{r}
quad %>% 
  mutate(glance = map(model, broom::glance))%>%
  unnest(glance, .drop = TRUE)%>%
  arrange(r.squared)
```
Compare to the linear model, the `r.squared` is much better.
```{r}
quad %>%
  mutate(glance = map(model, broom::glance))%>%
  unnest(glance, .drop = TRUE) %>%
  ggplot(aes(r.squared)) + 
  geom_histogram(bins = 100)
```

2. Explore other methods for visualising the distribution of `R^2` per continent. You might want to try the `ggbeeswarm` package, which provides similar methods for avoiding overlaps as jitter, but uses deterministic methods.
```{r}
library(ggbeeswarm)
gapminder %>%
  mutate(year = year - mean(year)) %>%
  group_by(continent,country)%>%
  nest() %>%
  mutate(model = map(data, modquad)) %>%
  mutate(glance = map(model, broom::glance))%>%
  unnest(glance) %>%
  ggplot(aes(x= continent, y =r.squared,color = continent)) + 
  geom_beeswarm()
```
It is clear that many Africa countries have patterns not captured by the polynomial functions.

3.To create the last plot (showing the data for the countries with the worst model fits), we needed two steps: we created a data frame with one row per country and then semi-joined it to the original dataset. It’s possible avoid this join if we use `nnest()` instead of `unnest(.drop = TRUE)`. How?

```{r}
gapminder %>%
  mutate(year = year - mean(year)) %>%
  group_by(country)%>%
  nest() %>%
  mutate(model = map(data, modquad)) %>%
  mutate(glance = map(model, broom::glance))%>%
  unnest(glance) %>%
  unnest(data) %>%
  semi_join(gapminder, by = c("pop","country")) %>%
  arrange(r.squared) %>%
  filter(r.squared %in% unique(r.squared)[1:6])%>%
  ggplot(aes(x = year + mean(gapminder$year), y = log(pop) )) +
  geom_line(aes(color = country))
```


### List-columns
