# Many models

Author: Ron 
Reviewer:
  
## Introduction

Note:

1. Use many simple models to better understand complex datasets
2. Use list-columns to store arbitrary data stracture in a data frame
3. Use **broom** package to turn models into tidy data

### Prerequisites
```{r}
library(modelr)
library(tidyverse)
library(gapminder)
```

## gapminder
```{r}
gapminder
```

### Nested data

use `nest()` function to create a dataframe of dataframes which has a column called `data`.

```{r}
by_country <- gapminder %>%
  group_by(country, continent) %>%
  nest()
```
Use `mutate` and `purrr::map` to generate new column, the model itself, which is an S3 object.

```{r}
country_model <- function(df) {
  lm(lifeExp ~ year, data = df)
}
by_country <- by_country %>% 
  mutate(model = map(data, country_model))
by_country
```

### Unnesting
`purrr::map2` can map mulitple inputs simultaneously
```{r}
by_country <- by_country %>%
  mutate(resids = map2(data, model, add_residuals)
         # data here is the column name, so is model.
         )

unnest(by_country,resids)
```
Note that each row is one *data point* in the original `gapminder` but the resid calculated is based on the country-wise model.

### Model quality

Use `broom` package to check model quality metrices. `broom::glance` takes a model as input

```{r}
glance <- by_country %>%
  mutate(glance = map(model, broom::glance)) %>%
  unnest(glance, .drop = TRUE)
# drop the list columns, `data`, `model` and `resid`
```
Note that each `model` gives a set of metrics.

```{r}
bad_fit <- filter(glance, r.squared < 0.25)

gapminder %>% 
  semi_join(bad_fit, by = "country") %>% 
  ggplot(aes(year, lifeExp, colour = country)) +
    geom_line()
```

### 25.2.5 Exercises

1. A linear trend seems to be slightly too simple for the overall trend. Can you do better with a quadratic polynomial? How can you interpret the coefficients of the quadratic? (Hint you might want to transform `year` so that it has mean zero.)

```{r}
modquad <- function(df){
  lm(data = df, lifeExp ~ poly(year,2))
}

quad <- gapminder %>%
  mutate(year = year - mean(year)) %>%
  group_by(country)%>%
  nest()%>%
  mutate(model = purrr::map(data,modquad))
  
```









