--- 
title: "Selected Solutions to R4DS Exercises"
author: "Chunji Wang"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: chunjiw/r4ds-solution
description: "This book provides selected solutions to the exercises in the wonderful book _R for Data Science_ by Wickham Hadley."
---

# Welcome {-}

This is the website for "**Selected Solutions to R4DS Exercises**". This is a joint advanture between [Chunji Wang](https://chunjiw.rbind.io), Ron, Luna, Zhiyin, Chengcheng.... We started the "**R4DS Study Club**" on Sep 22nd, 2017; If you want to join us, please contact [us](mailto:chunjiw@gmail.com)!

## How to use this book

The chapter labels in this book is the same as the original [R4DS](http://http://r4ds.had.co.nz) book; go to the corresponding chapter for solutions. You might need to read the beginning of the chapter to load some packages or create some variables that are needed in the exercises. 

## How to contribute to this book

The following are instructions of how to write a book with bookdown.

You can label chapter and section titles using `{#label}` after them, e.g., we can reference Chapter \@ref(intro). If you do not manually label them, there will be automatic labels anyway, e.g., Chapter \@ref(tibbles).

Figures and tables with captions will be placed in `figure` and `table` environments, respectively.

```{r nice-fig, fig.cap='Here is a nice figure!', out.width='80%', fig.asp=.75, fig.align='center'}
par(mar = c(4, 4, .1, .1))
plot(pressure, type = 'b', pch = 19)
```

Reference a figure by its code chunk label with the `fig:` prefix, e.g., see Figure \@ref(fig:nice-fig). Similarly, you can reference tables generated from `knitr::kable()`, e.g., see Table \@ref(tab:nice-tab).

```{r nice-tab, tidy=FALSE}
knitr::kable(
  head(iris, 20), caption = 'Here is a nice table!',
  booktabs = TRUE
)
```

You can write citations, too. For example, we are using the **bookdown** package [@R-bookdown] in this sample book, which was built on top of R Markdown and **knitr** [@xie2015].